---
output: pdf_document
---

# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

I load libraries.
```{r echo=TRUE , message=FALSE, warning=FALSE}
library("plyr")
library("ggplot2")
library("timeDate")
```

I load the data and I build a new data set without missing values(NA).
```{r echo=TRUE}
activity<-read.csv("data/activity.csv")
activitywithoutNa<-activity[!is.na(activity$steps),]
```
The data set look like this:
```{r echo=TRUE}
head(activity)

```

## What is mean total number of steps taken per day?
Firstly, I compute the total number of steps per day and then I make a histogram of the total number of steps taken each day. It show the distribution for this variable.
```{r echo=TRUE , message=FALSE, warning=FALSE}
totalStepDay<-ddply(activitywithoutNa,.(date),summarise,totalSteps=sum(steps))
m <- ggplot(totalStepDay, aes(x=totalSteps))
m + geom_histogram()
```
 
I compute the mean and the median of the total number of steps taken each day.
```{r echo=TRUE}
mean(totalStepDay$totalSteps)
median(totalStepDay$totalSteps)
```

## What is the average daily activity pattern?
Firstly, I compute the average of steps taken each interval and then I make time serie to show the average daily activity pattern.
```{r echo=TRUE}
averageStepInterval<-ddply(activitywithoutNa,.(interval),summarise,average=mean(steps))
m <- ggplot(averageStepInterval, aes(x=interval,y=average))
m + geom_line()
```

Here I show the interval with the highest number of steps.
```{r echo=TRUE}
averageStepInterval[which.max(averageStepInterval$average),1]
```

## Imputing missing values
Here I compute and report the total number of missing values in the dataset.
```{r echo=TRUE}
sum(is.na(activity$steps))
```
I use the mean of steps per interval to fill the missing values in the dataset and then I build a new dataset that is equal to the original dataset but with the missing data filled in.
```{r echo=TRUE}
activityCompleted<-join(activity, averageStepInterval, by = "interval", type = "left", match = "all")
activityCompleted[is.na(activity$steps),]$steps<-activityCompleted[is.na(activity$steps),4]
```

The new datset look like this:
```{r echo=TRUE}
head(activityCompleted)
```

With the new data set, I compute the total number of steps per day and then I make a histogram of the total number of steps taken each day. It show the distribution for this variable.
```{r echo=TRUE, message=FALSE, warning=FALSE}
totalStepDayAcCompleted<-ddply(activityCompleted,.(date),summarise,totalSteps=sum(steps))
m <- ggplot(totalStepDayAcCompleted, aes(x=totalSteps))
m + geom_histogram()
```

The mean and the meadia for total steps per day are:
```{r echo=TRUE}
mean(totalStepDayAcCompleted$totalSteps)
median(totalStepDayAcCompleted$totalSteps)
```
*The mean is equal to the estimate from the dataset without missing values and the median only differ in 1 with the estimates from the other data set.*


## Are there differences in activity patterns between weekdays and weekends?
Firstly, I build a new variable whit two values "weekend" and "weekday"
```{r}
activityCompleted$date<-as.Date(as.character(activityCompleted$date))
activityCompleted$partOfWeek[isWeekend(activityCompleted$date)]<-"weekend"
activityCompleted$partOfWeek[!isWeekend(activityCompleted$date)]<-"weekday"
activityCompleted$partOfWeek<-as.factor(activityCompleted$partOfWeek)
```
Then I compute the average of steps per interval on weekend days and also on weekday days. Moreover, I make time serie to show the average daily activity pattern on "weekend" and "weekday" 
```{r}
averageStepIntervalOfWeek<-ddply(activityCompleted,.(partOfWeek,interval),summarise,average=mean(steps))
m <- ggplot(averageStepIntervalOfWeek, aes(x=interval,y=average))
m + geom_line() + facet_grid(partOfWeek ~ . )
```


